Even-point Multi-loop Unitarity and its Applications: Exponentiation, Anomalies and Evanescence

### 2023-07-07 eventPointTwoLoop.tex
small changes
Final edits -- still need to check typos
made distinction between O vs T and added explicit operators definitions, O_i, that match the D-dimensional matrix elements, T_i. Added back in eBox (only for main results). Added a section on symmetri cdouble copy to gravitatioinal counter terms (and explained how we capture Boels' "exceptional" gravity amplitude -- and crush their PRL). Also added some extra citations for Jara's collaborator and Jake's recent double copy paper. Cut down on flowery language

\iffalse
\usepackage[framemethod=default]{mdframed}
\newmdenv[skipabove=7pt,
skipbelow=7pt,
rightline=false,
leftline=false,
topline=false,
bottomline=false,
backgroundcolor=gray!10,
linecolor=gray,
innerleftmargin=5pt,
innerrightmargin=5pt,
innertopmargin=5pt,
innerbottommargin=5pt,
leftmargin=0cm,
rightmargin=0cm,
linewidth=4pt]{eBox}
\fi

@@

\def\Ksl{{\s K}}
\def\ksl{\s{k}}
\def\Perm{{\cal P}}
\def\M{{\cal M}}
\def\ve{\varepsilon}
\def\tlambda{{\tilde\lambda}}
\def\MHVbar{$\overline{\hbox{MHV}}$}
\def\NMHVbar{$\overline{\hbox{NMHV}}$}
\def\P{{\rm P}}
\def\NHP{{\rm NHP}}
\def\mud{\lambda}
\def\bowtie{{\rm bow\mbox{-}tie}}
\newcommand{\cred}{\bf \color{red}}
 \newcommand{\cblue}{\color{blue}}
 \definecolor{MattOrange}{rgb}{1.0,0.4,0.2}
\newcommand{\cob}{ \bf \color{MattOrange}}
\newcommand{\andd}{\ , \quad \text{and}  \quad}
\newcommand{\forr}{\ , \quad \text{for}  \quad}

@@

In this paper we make the quantum leap to the multi-loop sector of effective field theory.  Namely we study even-point (EP) effective field theory (EFT) at the multi-loop level using a combination of generalized unitarity \cite{UnitarityMethod, Fusing, BDKUniarityReview} and the double-copy construction \cite{BCJ, BCJLoop}. The EFTs that we consider are the nonlinear sigma model (NLSM) and a related family of gauge theories involving Born-Infeld theory and supersymmetric generalizations known as Dirac-Born-Infeld-Volkov-Akulov (DBIVA) theories. These families are known at tree-level to be double-copies between NLSM and Yang-Mills theories with varying amounts of supersymmetry. While we focus on these specific examples, the methods we develop can be used for perturbative calculations in any even point effective field theory.

The motivation for this work is three-fold. First, perturbative calculations in NLSM and DBIVA effective field theories are in many ways significantly simpler than the typical multi-loop calculations of more phenomenological theories, like quantum-chromodynamics (QCD). As we will show, the multi-loop amplitudes for these theories allow us to recycle many of the $D$-dimensional integration tools that are so powerful at one-loop order. To approach this problem, we expand on the now standard approach of Forde \cite{Forde:2007mi}, that computes one-loop amplitudes directly from 4D unitarity cuts. In \sect{sec:EMU} we'll show that there are only two basis integrals needed at two-loop four-point for even point effective field theory amplitudes. In the spirit of one-loop unitarity, the integral coefficients can be directly extracted from EMU and tensor reduction. Our method of Even-point Multi-loop Unitarity (EMU) is a $D$-dimensional approach to integrand construction. This allows us to compute a large catalog of $D$-dimensional two-loop amplitudes, which provides fertile ground for cultivating insights about the multi-loop structure of effective field theories. The perturbative depth of our calculations sheds light on the exponential structure of IR divergences for NLSM amplitudes through two-loops, the anomalous behavior $U(1)$ duality invariance through two-loop order in DBIVA theories, as well as the emergence of evanescent operators relevant for anomaly cancellation in pure Born-Infeld theory. We believe these surprisingly rich physical structures could serve as a theoretical laboratory for future studies of the interplay between effective field theory operators and perturbative calculations in quantum field theory.

@@

The outline of the paper is as follows: in \sect{sec:Review}, we will provide a review of the on-shell methods and integration techniques needed to probe the multi-loop physics studied in this paper. Then in \sect{sec:EMU}, we introduce the method of Even-point Multi-loop Unitarity (EMU) and compute the two-loop tensor integrals needed in this work. In \sect{sec:Loops}, we present our results, beginning with a warm-up calculation in \sect{sec:NLSMU} where we construct the integrands for NLSM through NNLO in the effective coupling, and compute the fully integrated amplitudes. We show that in $D=2-2\epsilon$, the leading IR divergences of the theory exponentiate. With the scaffolding of $D$-dimensional integration in hand, we then compute two-loop amplitudes for $\mathcal{N}=4$ DBIVA theory and pure-photon Born-Infeld theory in \sect{sec:DBIU}. There we compute the anomalies present beyond one-loop order.  Then, using the NLSM integrands of \sect{sec:NLSMU}, we perform a multi-loop double-copy to $\mathcal{N}=4$ DBIVA observables with the color-dual basis numerators available in the literature in \sect{sec:DBIvDC}. After computing the catalog of two-loop amplitudes, we study the construction of quantum effective actions in \sect{sec:Actions} as they relate to the anomalous matrix elements present in two-loop Born-Infeld amplitudes. We demonstrate in \sect{sec:Anomalies} that the one-minus anomaly at two-loop requires the introduction of an evanescent operator at $\mathcal{O}(\alpha'^4)$. Then in \sect{sec:ActionDC} we take the opportunity to discuss the application of symmetric-structure double copy to cancel these anomalies, along with their relationship to higher-spin modes as studied in a recent work by the authors. Finally, in \sect{sec:EOpsCounting} we lay out a Hilbert series framework for counting evanescent operators at general orders in mass-dimension. To conclude, in \sect{sec:Discussion} we discuss many directions of future work and summarize the insights gained from this study.

@@

Throughout the text we will reserve $\mathcal{M}$ to denote double-copy amplitudes.  As shorthand, when the kinematic weights of two theories, $X$ and $Y$, participate in the double copy construction of $\mathcal{M}^{XY}$ we will use the outer-product to mean double copy construction as in,
\begin{equation}
\mathcal{M}^{\text{XY}} = X \otimes Y \equiv \int \prod_{i=1}^L\frac{d^D l_i}{(2\pi)^D} \sum_{g\in \Gamma^{(3)}_{n,L}} \frac{1}{S_g}\frac{N^X_gN^Y_g}{D_g}\,.
\end{equation}
At tree-level, such double-copy construction in the adjoint case can be understood equivalently \needRefs in terms of a KLT or momentum kernel matrix $S(a|b)$ and a BCJ spanning set of ordered amplitudes for each of th the $X$ and $Y$ theories,
\begin{equation}
 X \otimes Y = \sum_{a,b \in S_{n-3}(2,\ldots,n-2)} A^X(1,\{a\},n,n-1) S(a|b) A^Y(1,\{b\},n-1,n)\,,
\end{equation}
where the sum runs over both sets $a$, $b$ of $(n-3)!$ distinct ordered ampltidues.  No such momentum kernel has been constructed yet for the symmetric case.

@@

As noted, it will be convenient to take $D\rightarrow 4$ in certain circumstances after integration.  In such cases we will employ appropriate spinor-helicity variables.  Here we apply the same conventions of ref.~\cite{jjmcTASI2014}, which we quote now. For massless momenta $k_a$, and $k_b$ we have 
\begin{equation}
s_{ab} = (k_a+k_b)^2= \langle ab \rangle[ba]
\end{equation}
the component definition of our spinor bracket that are consistent with the conventions above,
\begin{align}
\langle ab \rangle &= \frac{(a_1 + i a_2)(b_0+b_3)-(b_1 + i b_2)(a_0+a_3)}{\sqrt{(a_0+a_3)(b_0+b_3)}}
\\
[ab] &= \frac{(b_1 - i b_2)(a_0+a_3)-(a_1 - i a_2)(b_0+b_3)}{\sqrt{(a_0+a_3)(b_0+b_3)}}
\end{align}
where the $a_i$ are component values of the four-vector, $k^\mu_a = (a_0,a_1,a_2,a_3)$.

 Four-dimensional polarization dot products with fixed helicity states can be mapped as follows:
\begin{equation}\label{eqn:4DPols}
\begin{aligned}
k_a \cdot \varepsilon_b^{(+)} &= \frac{\langle q a \rangle[ab]}{\sqrt{2}\langle q b\rangle}
\qquad\quad \qquad
k_a \cdot \varepsilon_b^{(-)} = -\frac{[qa]\langle ab\rangle}{\sqrt{2}[qb]}
\\
\varepsilon_a^{(-)}\cdot \varepsilon_b^{(+)} &= - \frac{\langle q a\rangle [qb]}{ [qa]\langle q b\rangle} 
\qquad \qquad
\varepsilon_a^{(\pm)}\cdot \varepsilon_b^{(\pm)} = 0 
\end{aligned}
\end{equation}
where $q^\mu$ is some null reference momentum that all polarizations are projected along.

While all four-dimensional equivalence relayed in this paper can be achieved analytically, in practice it is often much more  convenient to verify equivalence numerically.

@@

Furthermore, factors of $(\epsilon_i\cdot l)$ can be mapped to tensor structures with factors of $(\epsilon_i \cdot k_j)(k_i  \cdot l)/(k_i \cdot k_j)$ using polarization completeness relations~\cite{Bern:2017tuc} . Thus, any $N$-gon tensor integral permits a partial fraction decomposition in terms of inverse propagators, and thus can be mapped to a scalar basis of integrals:
\begin{equation}\label{TenRed1loop}
T_{\mu_1\mu_2...\mu_r} I_N^{\mu_1\mu_2...\mu_r} = \sum_{M=1}^N C_M I_M
\end{equation}
where $T_{\mu_1\mu_2...\mu_r}$ and $C_M$ are functions exclusively of external kinematics. This relationship holds $D$-dimensionally for any $r$-rank $N$-gon one-loop integral. Thus, any one-loop amplitude can be expressed completely in terms of a basis of one-loop integrals. This special property of one-loop integration is {not} the case for generic loop order, where the basis of ISPs grows faster than number of inverse propagators. As such, any universal integral basis beyond one-loop must include integrals with non-unit exponents for \dj{powers of} the propagators.

@@

The last important feature of one-loop basis integrals is that many of them can be evaluated in arbitrary dimension for a very general set of parameters appearing in the exponents of the propagators. \dj{Save Boxed for our key results?} Take for example the $D$-dimensional triangle and bubble integrals that we will use throughout the text:
\begin{equation}\label{eq:integrals}
{
\begin{aligned}
I_{2,(K)}^{(\alpha_1,\alpha_2)}&=\int \frac{d^{2D} l}{(2\pi)^{2D}} \frac{1}{[l^2]^{\alpha_1}[(l+K)^2]^{\alpha_2}} 
\\
&=i\left[-\frac{K^2}{4\pi}\right]^D\frac{\Gamma(D-\alpha_1)\Gamma(D-\alpha_2)\Gamma(\alpha_{12}-D)}{[K^2]^{\alpha_{12}}\Gamma(\alpha_1)\Gamma(\alpha_2)\Gamma(2D-\alpha_{12})} 
 \\\\
I_{3,(K_{12})}^{(\alpha_1,\alpha_2,\alpha_3)}&= \int \frac{d^{2D} l}{(2\pi)^{2D}} \frac{1}{[l^2]^{\alpha_1}[(l+K_1)^2]^{\alpha_2}[(l+K_{12})^2]^{\alpha_3}} 
\\
&=i\left[-\frac{K_{12}^2}{4\pi}\right]^D\frac{\Gamma(D-\alpha_{12})\Gamma(D-\alpha_{23})\Gamma(\alpha_{123}-D)}{[K_{12}^2]^{\alpha_{123}}\Gamma(\alpha_1)\Gamma(\alpha_3)\Gamma(2D-\alpha_{123})} 
\end{aligned}
}
\end{equation}
where we have introduced the notation $\alpha_{12...n} = \alpha_1+\alpha_2+\cdots +\alpha_n$, and have taken $D\rightarrow 2D$ to make the expressions more compact. The above integral expressions hold for massless external kinematics, $K_i^2=0$, which applies to all the integrals needed for the physical processes described in the text.

@@

Since the external states are massless, only the first two integrals, those appearing in \Fig{fig:contrib2Loop}, are non-vanishing in dimensional regularization -- the rest are scaleless. 
Therefore the above even-point perturbative predictions of $\varphi^{2k}$ is equivalent to $\varphi^4$ theory at four-point through two-loop order: 
\begin{equation}
\mathcal{M}_{\varphi^{2k}}^{\text{2-loop}} \equiv \mathcal{M}_{\varphi^{4}}^{\text{2-loop}} 
\end{equation}
In addition to being perturbatively equivalent to the quartic sector of the theory at two-loop, we can see above that the integrals for the Feynman diagrams in \eqn{eq:2loopEven} are {recursively one-loop}. In other words, to integrate the full amplitude at two-loop, we only need the iterated basis of one-loop integrals. This observation comes with the added advantage of permitting the application of one-loop Passarino-Veltman tensor reduction \cite{Passarino:1978jh} to two loop integrals. This property of recursively one-loop integrals will dramatically simplify the higher derivative even point theories that we consider in this work.

\begin{figure}[t]
    \centering
    \begin{equation*}\Omega_{(2,4)} = \left\{\scaleIntBscalar{}{}{}{},\scaleIntCscalar{}{}{}{} \right\}\end{equation*}
    \caption{Recursive one-loop integrals contributing to the  two-loop four-point calculation in even-point effective field theories, we refer to as the ostrich integral and double-bubble from left to right.}
    \label{fig:contrib2Loop}
\end{figure}

@@

We could also consider higher derivative generalizations of the NLSM that are invariant under the shift symmetry of \eqn{eq:shiftSym}.  Such  EFTs correspond to the so-called chiral limit\dj{Maybe cite Jara's collaborator -JJ} of {chiral perturbation theory}~\cite{Weinberg:1978kz,Gasser:1983yg}, describing massless pion scattering below the chiral symmetry breaking scale of QCD. 
\begin{equation}\label{eq:XPTLag}
\mathcal{L}^{\chi\text{PT}}=\frac{1}{2}\text{tr}[(\partial_\mu U)^\dagger (\partial ^\mu U)] + \frac{\beta_1}{f_\pi^2} \text{tr}[(\partial_\mu U)^\dagger (\partial ^\mu U)]^2 + \frac{\beta_2}{f_\pi^2} \text{tr}[\partial_\mu U^\dagger \partial_\nu U]\text{tr}[\partial_\mu U^\dagger \partial_\nu U] +\cdots
\end{equation}
By construction, each of the higher derivative operators appearing above are invariant under the shift symmetry of \eqn{eq:shiftSym}. These higher derivative generalizations of NLSM have been studied both in the context of the $S$-matrix bootstrap \cite{Manohar:2008tc,Bellazzini:2016xrt,Guerrieri:2020bto}, and soft theorem bootstraps of refs. \cite{Cheung:2014dqa,Cheung:2015ota,Cheung:2016drk,Cheung:2018oki,Low:2019ynd}. Considering the relevance of \eqn{eq:XPTLag} in Standard Model pion scattering, the amplitudes for $SU(2)$ NLSM have been computed through two-loop order \cite{Bijnens:1995yn,Bijnens:1997vq,Girlanda:1997ed} . In this work, we add to the literature by computing the two-loop amplitudes of \eqn{eq:NLSMLag} for arbitrary $SU(N_c)$ color structure. In doing so, we show that the IR divergences of the theory exponentiates in the large $N_c$ limit. 
\paragraph{Born-Infeld Photons and Dirac Scalars}
The second even-point EFT we will study is Born-Infeld theory \cite{Born:1934gh}. This theory describes the dynamics of open string endpoints with $U(1)$ charge attached to a $D$-dimensional spacetime with Dirichlet boundary conditions. Given the stringy dynamics orthogonal to the spacetime, the electric field generated by the charged endpoints has a maximum field strength of order the inverse string tension, $1/\alpha'$. The effective Lagrangian for this theory
\begin{equation}
 \label{biLag}
  \alpha'^{\,2}\mathcal{L}^{\text{BI}} = 1-\sqrt{\text{det}(\eta_{\mu\nu}+\alpha' F_{\mu\nu})} \,.
\end{equation}
A close cousin of Born-Infeld theory is the dimensionally reduction to Dirac scalars, which describes a $D$-dimensional spacetime propagating in a $(D+1)$-dimensional background:
\begin{equation}
 \label{DBI}
  \alpha'^{\,2}\mathcal{L}^{\text{DBI}} =1- \sqrt{\text{det}(\eta_{\mu\nu}+\alpha'^{\,2} \partial_\mu \varphi \partial_\nu \varphi)}
\end{equation}
where $\varphi$ is the spacetime coordinate in the orthogonal space. This theory is uniquely determined by considering the most general polynomial  of $X=(\partial \varphi)^2$, $P(X)$,  constrained to be  invariant under the field redefinition~ \cite{Cheung:2016drk}:
\begin{equation}\label{DBIShift}
\varphi \rightarrow \varphi + c + b^{\mu} (x_\mu + \varphi \partial_\mu \varphi)
\end{equation}

@@

It is worth noting that in the case of supersymmetric DBIVA, at four-points, $U(1)$ symmetry is promoted to an $R$-symmetry, and is protected perturbatively to all orders by supersymmetric Ward Identities \cite{Heydeman:2017yww}. We will show this explicitly in the case of $\mathcal{N}=4$ DBIVA through two-loop. However, we note that this supersymmetric enhancement of $U(1)$ duality does not carry over to higher multiplicity, where $R$-symmetry permits non-vanishing matrix elements outside the split helicity sector. To see one such example of this, we need not look any further than the abelianized OSS described above. At higher-multiplicity and higher orders in $\alpha'$, the manifestly supersymmetric  abelian OSS violates $U(1)$ duality. It would be fascinating if these higher derivative violations of $U(1)$ duality in OSS spectrum are precisely those needed to cancel anomalous behavior at the quantum level of higher multiplicity amplitudes. \dj{Maybe some more words here?  I'm not sure what we're saying -- are we talking about stuff required to eat up anomalous 6-point behavior?} We see this as a natural future direction and application of the methods we develop in this work.

>>>>

\usepackage[framemethod=default]{mdframed}
\newmdenv[skipabove=7pt,
skipbelow=7pt,
rightline=false,
leftline=false,
topline=false,
bottomline=false,
backgroundcolor=gray!10,
linecolor=gray,
innerleftmargin=5pt,
innerrightmargin=5pt,
innertopmargin=5pt,
innerbottommargin=5pt,
leftmargin=0cm,
rightmargin=0cm,
linewidth=4pt]{eBox}

@@

\def\Ksl{{\s K}}
\def\ksl{\s{k}}
\def\Perm{{\cal P}}
\def\M{{\cal M}}
\def\ve{\varepsilon}
\def\tlambda{{\tilde\lambda}}
\def\MHVbar{$\overline{\hbox{MHV}}$}
\def\NMHVbar{$\overline{\hbox{NMHV}}$}
\def\P{{\rm P}}
\def\NHP{{\rm NHP}}
\def\mud{\lambda}
\def\bowtie{{\rm bow\mbox{-}tie}}
\newcommand{\cred}{\bf \color{red}}
 \newcommand{\cblue}{\color{blue}}
\newcommand{\andd}{\ , \quad \text{and}  \quad}
\newcommand{\forr}{\ , \quad \text{for}  \quad}

@@

In this paper we make the quantum leap to the multi-loop sector of effective field theory.  Namely we study even-point (EP) effective field theory (EFT) at the multi-loop level using a combination of generalized unitarity \cite{UnitarityMethod, Fusing, BDKUniarityReview} and the double-copy construction \cite{BCJ, BCJLoop}. The EFTs that we consider are the nonlinear sigma model (NLSM) and a related family of gauge theories involving Born-Infeld theory and supersymmetric generalizations known as Dirac-Born-Infeld-Volkov-Akulov (DBIVA) theories. These families are known at tree-level to be double-copies between NLSM and Yang-Mills theories with varying amounts of supersymmetry. While we focus on these specific examples, the methods we develop can be used for perturbative calculations in any even-point effective field theory.

The motivation for this work is three-fold. First, perturbative calculations in NLSM and DBIVA effective field theories are in many ways significantly simpler than the typical multi-loop calculations of more phenomenological theories, like quantum-chromodynamics (QCD). As we will show, the multi-loop amplitudes for these theories allow us to recycle many of the $D$-dimensional integration tools that are so powerful at one-loop order. To approach this problem, we expand on the now standard procedure of Forde \cite{Forde:2007mi}, that computes one-loop amplitudes directly from 4D unitarity cuts. In \sect{sec:EMU} we'll show that there are only two basis integrals needed at two-loop four-point for even-point effective field theory amplitudes. In the spirit of one-loop unitarity, the integral coefficients can be directly extracted from Even-point Multi-loop Unitarity and tensor reduction. Our method of Even-point Multi-loop Unitarity (EMU) is a $D$-dimensional approach to integrand construction. This allows us to compute a large catalog of $D$-dimensional two-loop amplitudes, which provides fertile ground for cultivating insights about the multi-loop structure of effective field theories. The perturbative depth of our calculations sheds light on the exponential structure of IR divergences for NLSM amplitudes through two-loops, the anomalous behavior $U(1)$ duality invariance through two-loop order in DBIVA theories, as well as the emergence of evanescent operators relevant for anomaly cancellation in pure Born-Infeld theory. We believe these surprisingly rich physical structures could serve as a theoretical laboratory for future studies of the interplay between effective field theory operators and perturbative calculations in quantum field theory.

@@

The outline of the paper is as follows: in \sect{sec:Review}, we will provide a review of the on-shell methods and integration techniques needed to probe the multi-loop physics studied in this paper. Then in \sect{sec:EMU}, we introduce the method of Even-point Multi-loop Unitarity (EMU) and compute the two-loop tensor integrals needed in this work. In \sect{sec:Loops}, we present our results, beginning with a warm-up calculation in \sect{sec:NLSMU} where we construct integrands for NLSM through two-loop order, and compute the fully integrated amplitudes. We show that in $D=2-2\epsilon$, the leading IR divergences of the theory exponentiate. With the scaffolding of $D$-dimensional integration in hand, we then compute two-loop amplitudes for $\mathcal{N}=4$ DBIVA theory and pure-photon Born-Infeld theory in \sect{sec:DBIU}. There we compute the anomalies present beyond one-loop order.  Then, using the NLSM integrands of \sect{sec:NLSMU}, we perform a multi-loop double-copy to $\mathcal{N}=4$ DBIVA observables with the color-dual basis numerators available in the literature in \sect{sec:DBIvDC}. After computing the catalog of two-loop amplitudes, we study the construction of quantum effective actions in \sect{sec:Actions} as they relate to the anomalous matrix elements present in two-loop Born-Infeld amplitudes. We demonstrate in \sect{sec:Anomalies} that the one-minus anomaly at two-loop requires the introduction of an evanescent operator at $\mathcal{O}(\alpha'^4)$. Then in \sect{sec:ActionDC} we take the opportunity to discuss the application of symmetric-structure double copy to cancel these anomalies, along with their relationship to higher-spin modes as studied in a recent work by the authors. Finally, in \sect{sec:EOpsCounting} we lay out a Hilbert series framework for counting evanescent operators at general orders in mass-dimension. To conclude, in \sect{sec:Discussion} we discuss many directions of future work and summarize the insights gained from this study.

@@

Throughout the text we will reserve $\mathcal{M}$ to denote double-copy amplitudes.  As shorthand, when the kinematic weights of two theories, $X$ and $Y$, participate in the double copy construction of $\mathcal{M}^{XY}$ we will use the outer-product to mean double copy construction as in,
\begin{equation}
\mathcal{M}^{\text{XY}} = X \otimes Y \equiv \int \prod_{i=1}^L\frac{d^D l_i}{(2\pi)^D} \sum_{g\in \Gamma^{(3)}_{n,L}} \frac{1}{S_g}\frac{N^X_gN^Y_g}{D_g}\,.
\end{equation}
At tree-level, such double-copy construction in the adjoint case can be understood equivalently \cite{Bern:1998sv,Vaman:2010ez,Bjerrum-Bohr:2010diw,Bjerrum-Bohr:2010kyi,Bjerrum-Bohr:2010pnr,Bjerrum-Bohr:2012kaa} in terms of a KLT or momentum kernel matrix $S(a|b)$ and a BCJ spanning set of ordered amplitudes for each of th the $X$ and $Y$ theories,
\begin{equation}
 X \otimes Y = \sum_{a,b \in S_{n-3}(2,\ldots,n-2)} A^X(1,\{a\},n,n-1) S(a|b) A^Y(1,\{b\},n-1,n)\,,
\end{equation}
where the sum runs over both sets $a$, $b$ of $(n-3)!$ distinct ordered ampltidues.  No such momentum kernel has been constructed yet for the symmetric case.

@@

As noted, it will be convenient to take $D\rightarrow 4$ in certain circumstances after integration.  In such cases we will employ appropriate spinor-helicity variables.  Here we apply the same conventions of ref.~\cite{jjmcTASI2014}, which we quote now. For massless momenta $k_a$, and $k_b$ we have 
\begin{equation}
s_{ab} = (k_a+k_b)^2= \langle ab \rangle[ba]
\end{equation}
the component definition of our spinor bracket that are consistent with the conventions above,
\begin{align}
\langle ab \rangle &= \frac{(a_1 + i a_2)(b_0+b_3)-(b_1 + i b_2)(a_0+a_3)}{\sqrt{(a_0+a_3)(b_0+b_3)}}
\\
[ab] &= \frac{(b_1 - i b_2)(a_0+a_3)-(a_1 - i a_2)(b_0+b_3)}{\sqrt{(a_0+a_3)(b_0+b_3)}}
\end{align}
where the $a_i$ are component values of the four-vector, $k^\mu_a = (a_0,a_1,a_2,a_3)$. Four-dimensional polarization dot products with fixed helicity states can be mapped as follows:
\begin{equation}\label{eqn:4DPols}
\begin{aligned}
k_a \cdot \varepsilon_b^{(+)} &= \frac{\langle q a \rangle[ab]}{\sqrt{2}\langle q b\rangle}
\qquad\quad \qquad
k_a \cdot \varepsilon_b^{(-)} = -\frac{[qa]\langle ab\rangle}{\sqrt{2}[qb]}
\\
\varepsilon_a^{(-)}\cdot \varepsilon_b^{(+)} &= - \frac{\langle q a\rangle [qb]}{ [qa]\langle q b\rangle} 
\qquad \qquad
\varepsilon_a^{(\pm)}\cdot \varepsilon_b^{(\pm)} = 0 
\end{aligned}
\end{equation}
where $q^\mu$ is some null reference momentum that all polarizations are projected along. While all four-dimensional equivalence relayed in this paper can be achieved analytically, in practice it is often much more  convenient to verify equivalence numerically.

@@

Furthermore, factors of $(\varepsilon_i\cdot l)$ can be mapped to tensor structures of external kinematics with factors of $(\varepsilon_i \cdot k_j)(k_i  \cdot l)/(k_i \cdot k_j)$ using polarization completeness relations~\cite{Bern:2017tuc} . Thus, any $N$-gon tensor integral permits a partial fraction decomposition in terms of inverse propagators, and thus can be mapped to a scalar basis of integrals:
\begin{equation}\label{TenRed1loop}
T_{\mu_1\mu_2...\mu_r} I_N^{\mu_1\mu_2...\mu_r} = \sum_{M=1}^N C_M I_M
\end{equation}
where $T_{\mu_1\mu_2...\mu_r}$ and $C_M$ are functions exclusively of external kinematics. This relationship holds $D$-dimensionally for any $r$-rank $N$-gon one-loop integral. Thus, any one-loop amplitude can be expressed completely in terms of a basis of one-loop integrals. This special property of one-loop integration is {not} the case for generic loop order, where the basis of ISPs grows faster than number of inverse propagators. As such, any universal integral basis beyond one-loop must include integrals with non-unit powers of the propagators.

@@

The last important feature of one-loop basis integrals is that many of them can be evaluated in arbitrary dimension for a very general set of parameters appearing in the exponents of the propagators. Take for example the $D$-dimensional triangle and bubble integrals \cite{Smirnov:2004ym} that we will use throughout the text:
\begin{equation}\label{eq:integrals}
{
\begin{aligned}
I_{2,(K)}^{(\alpha_1,\alpha_2)}&=\int \frac{d^{2D} l}{(2\pi)^{2D}} \frac{1}{[l^2]^{\alpha_1}[(l+K)^2]^{\alpha_2}} 
\\
&=i\left[-\frac{K^2}{4\pi}\right]^D\frac{\Gamma(D-\alpha_1)\Gamma(D-\alpha_2)\Gamma(\alpha_{12}-D)}{[K^2]^{\alpha_{12}}\Gamma(\alpha_1)\Gamma(\alpha_2)\Gamma(2D-\alpha_{12})} 
 \\\\
I_{3,(K_{12})}^{(\alpha_1,\alpha_2,\alpha_3)}&= \int \frac{d^{2D} l}{(2\pi)^{2D}} \frac{1}{[l^2]^{\alpha_1}[(l+K_1)^2]^{\alpha_2}[(l+K_{12})^2]^{\alpha_3}} 
\\
&=i\left[-\frac{K_{12}^2}{4\pi}\right]^D\frac{\Gamma(D-\alpha_{12})\Gamma(D-\alpha_{23})\Gamma(\alpha_{123}-D)}{[K_{12}^2]^{\alpha_{123}}\Gamma(\alpha_1)\Gamma(\alpha_3)\Gamma(2D-\alpha_{123})} 
\end{aligned}
}
\end{equation}
where we have introduced the notation $\alpha_{12...n} = \alpha_1+\alpha_2+\cdots +\alpha_n$, and have taken $D\rightarrow 2D$ to make the expressions more compact. The above integral expressions hold for massless external kinematics, $K_i^2=0$, which applies to all the integrals needed for the physical processes described in the text.

@@

Since the external states are massless, only the first two integrals, those appearing in \Fig{fig:contrib2Loop}, are non-vanishing in dimensional regularization -- the rest are scaleless. 
Therefore the above even-point perturbative predictions of $\varphi^{2k}$ is equivalent to $\varphi^4$ theory at four-point through two-loop order: 
\begin{equation}
\mathcal{M}_{\varphi^{2k}}^{\text{2-loop}} \equiv \mathcal{M}_{\varphi^{4}}^{\text{2-loop}} 
\end{equation}
In addition to being perturbatively equivalent to the quartic sector of the theory at two-loop, we can see above that the integrals for the Feynman diagrams in \eqn{eq:2loopEven} are {recursively one-loop}. In other words, to integrate the full amplitude at two-loop, we only need the iterated basis of one-loop integrals. This observation comes with the added advantage of permitting the application of one-loop Passarino-Veltman tensor reduction \cite{Passarino:1978jh} to two loop integrals. This property of recursively one-loop integrals will dramatically simplify the higher derivative even-point theories that we consider in this work.

\begin{figure}[t]
    \centering
    \begin{equation*}\Omega_{(2,4)} = \left\{\scaleIntBscalar{}{}{}{},\scaleIntCscalar{}{}{}{} \right\}\end{equation*}
    \caption{Recursive one-loop integrals contributing to the  two-loop four-point calculation in even-point effective field theories, we refer to as the peacock integral and double-bubble from left to right.}
    \label{fig:contrib2Loop}
\end{figure}

@@

We could also consider higher derivative generalizations of the NLSM that are invariant under the shift symmetry of \eqn{eq:shiftSym}.  Such  EFTs correspond to the so-called chiral limit of {chiral perturbation theory}~\cite{Weinberg:1978kz,Gasser:1983yg,Kampf:2006yf}, describing massless pion scattering below the chiral symmetry breaking scale of QCD:
\begin{equation}\label{eq:XPTLag}
\mathcal{L}^{\chi\text{PT}}=\frac{1}{2}\text{tr}[(\partial_\mu U)^\dagger (\partial ^\mu U)] + \frac{\beta_1}{f_\pi^2} \text{tr}[(\partial_\mu U)^\dagger (\partial ^\mu U)]^2 + \frac{\beta_2}{f_\pi^2} \text{tr}[\partial_\mu U^\dagger \partial_\nu U]\text{tr}[\partial_\mu U^\dagger \partial_\nu U] +\cdots
\end{equation}
By construction, each of the higher derivative operators appearing above are invariant under the shift symmetry of \eqn{eq:shiftSym}. These higher derivative generalizations of NLSM have been studied both in the context of the $S$-matrix bootstrap \cite{Manohar:2008tc,Bellazzini:2016xrt,Guerrieri:2020bto}, and soft theorem bootstraps of refs. \cite{Cheung:2014dqa,Cheung:2015ota,Cheung:2016drk,Cheung:2018oki,Low:2019ynd}. Considering the relevance of \eqn{eq:XPTLag} in Standard Model pion scattering, the amplitudes for $SU(2)$ NLSM have been computed through two-loop order \cite{Bijnens:1995yn,Bijnens:1997vq,Girlanda:1997ed}. In this work, we add to the literature by computing the two-loop amplitudes of \eqn{eq:NLSMLag} for arbitrary $SU(N_c)$ color structure. In doing so, we show that the IR divergences of the theory exponentiates in the large $N_c$ limit. 
\paragraph{Born-Infeld Photons and Dirac Scalars}
The second even-point EFT we will study is Born-Infeld theory \cite{Born:1934gh}. This theory describes the dynamics of open string endpoints with $U(1)$ charge attached to a $D$-dimensional spacetime with Dirichlet boundary conditions. Given the stringy dynamics orthogonal to the spacetime, the electric field generated by the charged endpoints has a maximum field strength of order the inverse string tension, $1/\alpha'$. The effective Lagrangian for this theory
\begin{equation}
 \label{biLag}
  \alpha'^{\,2}\mathcal{L}^{\text{BI}} = 1-\sqrt{\text{det}(\eta_{\mu\nu}+\alpha' F_{\mu\nu})} \,.
\end{equation}
A close cousin of Born-Infeld theory is the dimensionally reduction to Dirac scalars, which describes a $D$-dimensional spacetime propagating in a $(D+1)$-dimensional background:
\begin{equation}
 \label{DBI}
  \alpha'^{\,2}\mathcal{L}^{\text{DBI}} =1- \sqrt{\text{det}(\eta_{\mu\nu}+\alpha'^{\,2} \partial_\mu \varphi \partial_\nu \varphi)}
\end{equation}
where $\varphi$ is the spacetime coordinate in the orthogonal space. This theory is uniquely determined by considering the most general polynomial  of $X=(\partial \varphi)^2$, $P(X)$,  constrained to be  invariant under the field redefinition~ \cite{Cheung:2016drk}:
\begin{equation}\label{DBIShift}
\varphi \rightarrow \varphi + c + b^{\mu} (x_\mu + \varphi \partial_\mu \varphi)
\end{equation}

@@

Before proceeding, we comment briefly on the higher derivative corrections one might expect to appear in duality invariant effective field theories. In the case of supersymmetric DBIVA at four-points, $U(1)$ symmetry is promoted to an $R$-symmetry, and is thus protected perturbatively to all orders by supersymmetric Ward Identities \cite{Heydeman:2017yww}. We show this explicitly in the case of $\mathcal{N}=4$ DBIVA through two-loop in \sect{sec:DBIU}. However, this supersymmetric enhancement of $U(1)$ duality does not in general apply to higher multiplicity amplitudes. Beyond four-point $R$-symmetry permits non-vanishing matrix elements outside the split helicity sector, beyond leading order in the higher-derivative expansion \cite{Pavao:2022kog}. The low energy effective action of the OSS is an exemplar of an EFT that respects U(1) duality at leading order in $\alpha'$, but permits duality violating matrix elements at six-point and beyond \cite{Carrasco:2016ldy}. At these higher orders in $\alpha'$, while preserving $R$-symmetry, the abelian OSS violates $U(1)$ duality at tree-level. It would be fascinating if these higher derivative violations of $U(1)$ duality in OSS spectrum are precisely those needed to cancel anomalous behavior at the quantum level. We see this as a natural future direction and application of the methods we develop in this work.
